Hierarchical Inference of the Intrinsic Ellipticity Distribution of Galaxies during Weak Lensing Analysis

### 2013-04-13 docs/ms.tex
Update ms.tex

Since $p(\alpha)$ does not explicitly depend on $\vec{g}$. $p(\alpha|\vec{g})$ is
just the prior distribution $p(\vec{\alpha}|\xi)$. Therefore, we have
marginalized likehood 
\begin{equation}
p(\epsilon_{\ell}|\vec{g})=\int p(\epsilon_{\ell}|\vec{g},\vec{\alpha})\, p(\vec{\alpha}|\xi)\,d\vec{\alpha}
\end{equation}
Given that marginalized likelihood function, in principle we could optimize it and gives out a max-likelihood estimator for $\vec{g}$
in different patches simultaneously.

@@

The correctness of posterior distribution of g depends on the intrinsic
ellipticity distribution P($\epsilon_{0}|\alpha$). Due to the limitation
of human's knowledge we will not be able to give an analytical form for
$P(\epsilon_{0}|\alpha)$. Therefore we proposed an exceedingly flexible
model for $P(\epsilon_{0}|\alpha)$, which does not requires further
assumption on $P(\epsilon_{0}|\alpha)$.

We divide the 2-D space r $\in$(0,1) into N bins. The probability
of each bin is $p_{i}=e^{\alpha_{i}}$. $\overrightarrow{\alpha}$
is a N dimension parameter to be infered, simultaneously with $g$.

@@

One thing to note is that the conditional probability of g in one patch only explicitly depends on $\alpha$, but not g in other patches. Therefore
with $\vec{\alpha}$ fixed, the posterior distributions of reduced shear in different patches are independent from each other. 
\begin{equation}
p(\vec{g}_{1}....\vec{g}_{n}|\vec{\alpha},Data)=p(\vec{g}_{1}|\vec{\alpha},Data_{1})....p(\vec{g}_{n}|\vec{\alpha},Data_{n})
\end{equation}
Therefore sampling reduced shear g from the conditional probability distribution can be reduced to a 2D sampling problem, which can be done very quickly.
In the Gibbs sampler, first we make a sample of $p(\vec{\alpha}|\vec{g},Data)$ with Metropolis-Hasting sampler. The we make samples of $g$ in different patches
independently according to formula (14).
The total time needed to sample $g$ and $\alpha$ of 64 sky patches with 8 galaxies in each patch will only be 
approximitly 25s on a 2 GHz CPU.

@@

The probability distribution can be visualised by plotting points at each sample
position. An example in the x, y plane is shown for the posterior distribution of
$\vec{g}$ of 64 sky patches in Figure 1. There are 8 galaxies in each patch in the simulation.
The number density of samples is proportional to the probability density. The figure shows that the two parameters of $\vec{g}$
do not have strong correlation, since the shape of smaples is close to a circle.
We also shows the histogram of g1 in a single sky patch in Figure 2. Not surprisingly it is close to a Gaussian distribution.
If the probability distribution of parameters are perfect Gaussian, the most probable estimates and uncertainty 
is just the mean and standard deviation of the Gaussian distribution. However, in pratical, we find that the distribution 
of reduced shear $\vec{g}$ will  get closer to Gaussian, when larger number of galaxies in one patch  is used. When the number of galaxies in 
one patch is equal to or larger than 8, we found it's eligible to make such an approximation that to use sample mean and standard 
deviation as the most probable estimates and uncertainty. We'll justify this approximation with a bias test in the following
sub section.

@@

\caption{The figure shows the estimated $g_{1}$ values minus the true (input)
$g_{1}^{true}$. The distribution range of $g_{1}^{true}$ is from
-0.1 to 0.1. Originally there are 640,000 true $g_{1}$. We bin the
plot for presentation purpose. The data point and error bars are the weighted
 average and  variance of $\delta g_{1}$ in each bin. The blue
line is the result of weighted linear fit of the original 640,000 data points
with sample standard deviation as error bar.}
\end{figure}
Suppose we have perfect measurements on galaxy shape. The measured ellipticity then equals to the true lensed ellipticity. With equation (1) the mean ellipticity
estimator has following expression :
\begin{equation}
<\epsilon_{\ell}>=g^{t}+<\epsilon_{0}>+<(\epsilon_{0}+g^{t})[-g^{t*}\epsilon_{0}+O(g^{t*}\epsilon_{0})]>
\end{equation}
When the unlensd ellipticity is isotropic ,we have $<\epsilon_{\ell}>=g^{t}$ . In this case, the mean ellipticity estimator is 
unbiased.
Meanwhile we find that simply using the posterior mean as the estimator of g will lead to a significant bias. 
Figure 3 shows the result of bias test for both mean ellipticity and posterior mean estimator.

@@

\subsection{Measurement Error}
In this section we  discuss how the shape measurement error 
affect the variance and bias of this Hierarchical Bayesian method.
In principle, if the likelihood of true lensed ellipticity $p(\epsilon_{ob}|\epsilon_{\ell})$
is known the likelihood $p(\epsilon_{ob}|g,\alpha)$ can be calculated using the convolution as equation ().
However, due to the limitation of computational time we avoid to do this convolution but use $\epsilon_{ob}$
as an estimator of $\epsilon_{\ell}$.
The likelihood of true lensed ellipticity $p(\epsilon_{ob}|\epsilon_{\ell})$ we use in the simulation is a 2D
Gaussian with mean equals $\epsilon_{\ell}$ and $\sigma$ equals 0.05 .
The result of bias test and mean square error test is shown in

>>>>

Since $p(\alpha)$ does not explicitly depend on $\vec{g}$. $p(\alpha|\vec{g})$ is
just the prior distribution $p(\vec{\alpha}|\xi)$. Therefore, we have
marginalized likehood 
\begin{equation}
p(\epsilon_{\ell}|\vec{g})=\int p(\epsilon_{\ell}|\vec{g},\vec{\alpha})\, p(\vec{\alpha}|\xi)\,d\vec{\alpha}
\end{equation}
Given that marginalized likelihood function, in principle we could optimize it and give out a max-likelihood estimator for $g$
in different patches simultaneously.

@@

The correctness of posterior distribution of $g$ depends on the intrinsic
ellipticity distribution P($\epsilon_{0}|\alpha$). Due to the limitation
of human's knowledge we will not be able to give an analytical form for
$P(\epsilon_{0}|\alpha)$. Therefore we propose an exceedingly flexible
model for $P(\epsilon_{0}|\alpha)$, which does not require further
assumption on $P(\epsilon_{0}|\alpha)$.

We divide the 2-D space r $\in$(0,1) into N bins. The probability
of each bin is $p_{i}=e^{\alpha_{i}}$. $\overrightarrow{\alpha}$
is a N dimension parameter to be inferred, simultaneously with $g$.

@@

One thing to note is that the conditional probability of $g$ in one patch only explicitly depends on $\alpha$, but not g in other patches. Therefore
with $\vec{\alpha}$ fixed, the posterior distributions of reduced shear in different patches are independent from each other. 
\begin{equation}
p(\vec{g}_{1}....\vec{g}_{n}|\vec{\alpha},Data)=p(\vec{g}_{1}|\vec{\alpha},Data_{1})....p(\vec{g}_{n}|\vec{\alpha},Data_{n})
\end{equation}
Therefore sampling reduced shear g from the conditional probability distribution can be reduced to a 2D sampling problem, which can be done very quickly.
In the Gibbs sampler, first we make a sample of $p(\vec{\alpha}|\vec{g},Data)$ with Metropolis-Hasting sampler. The we make samples of $g$ in different patches
independently according to formula (14).
The total time needed to sample $g$ and $\alpha$ of 64 sky patches with 8 galaxies in each patch is 
approximitly 25s on a 2 GHz CPU.

@@

The probability distribution can be visualised by plotting points at each sample
position. An example in the x, y plane is shown for the posterior distribution of
$g$ in 64 sky patches in Figure 1. There are 8 galaxies in each patch in the simulation.
The number density of samples is proportional to the probability density. The figure shows that the two parameters of $\vec{g}$
do not have strong correlation, since the shape of smaples is close to a circle.
We also shows the histogram of g1 in a single sky patch in Figure 2. Not surprisingly it is close to a Gaussian distribution.
If the probability distribution of parameters are perfect Gaussian, the most probable estimates and uncertainty 
is just the mean and standard deviation of the Gaussian distribution. However, in pratical, we find that the distribution 
of reduced shear $\vec{g}$ will  get closer to Gaussian, when larger number of galaxies in one patch  is used. When the number of galaxies in 
one patch is equal to or larger than 8, we found it's eligible to make such an approximation that to use sample mean and standard 
deviation as the most probable estimates and uncertainty. We'll justify this approximation with a bias test in the following
sub section.

@@

\caption{The figure shows the estimated $g_{1}$ values minus the true (input)
$g_{1}^{true}$. The distribution range of $g_{1}^{true}$ is from
-0.1 to 0.1. Originally there are 640,000 true $g_{1}$. We bin the
plot for presentation purpose. The data point and error bars are the weighted
 average and  variance of $\delta g_{1}$ in each bin. The blue
line is the result of weighted linear fit of the original 640,000 data points
with sample standard deviation as error bar.}
\end{figure}
Suppose we have perfect measurements on galaxy shape. The measured ellipticity then equals to the true lensed ellipticity. With equation (1) the mean ellipticity
estimator has following expression :
\begin{equation}
<\epsilon_{\ell}>=g^{t}+<\epsilon_{0}>+<(\epsilon_{0}+g^{t})[-g^{t*}\epsilon_{0}+O(g^{t*}\epsilon_{0})]>
\end{equation}
When the unlensd ellipticity is isotropic ,we have $<\epsilon_{\ell}>=g^{t}$ . In this case, the mean ellipticity estimator is 
unbiased.
Meanwhile we find that simply using the posterior mean as the estimator of $g$ will lead to a significant bias. 
Figure 3 shows the result of bias test for both mean ellipticity and posterior mean estimator.

@@

\subsection{Measurement Error}
In this section we  discuss how the shape measurement error 
affects the variance and bias shear inference.
In principle, if the likelihood of true lensed ellipticity $p(\epsilon_{ob}|\epsilon_{\ell})$
is known the likelihood $p(\epsilon_{ob}|g,\alpha)$ can be calculated using the convolution as equation ().
However, due to the limitation of computational time we avoid to do this convolution but use $\epsilon_{ob}$
as an estimator of $\epsilon_{\ell}$.
The likelihood of true lensed ellipticity $p(\epsilon_{ob}|\epsilon_{\ell})$ we use in the simulation is a 2D
Gaussian with mean equals $\epsilon_{\ell}$ and $\sigma$ equals 0.05 .
The result of bias test and mean square error test is shown in
